Designing and Developing a Personalised Recommender System

### 2021-02-11 paper/paper.tex
added lots of equations to algorithm section

\subsection{Related work review}
Hybrid recommender systems which combine Knowledge-based (KB) and Collaborative-filtering (CF) techniques 
are designed to combine the ability of KB techniques to give generic recommendations to new users with 
small profiles and the strength of CF to find peer users with unexpected shared preferences. 
This combination of techniques is particularly well suited to restaurants since a user may often be in the 
mood for a specific type of food, such as Italian or Chinese, so they only want to see suggestions of that 
type or restaurant, which is where the Knowledge-based recommender is helpful. 
A Collaborative-filtering algorithm will then be able to order all the restaurants which match this 
explicit criteria by how the user should rate them, based on the opinions of other similar users, 
helping them to discover new restaurants to try. 
These techniques are used in the Entree restaurant recommender system \cite{burke1999integrating,burke2000knowledge,burke2002hybrid,burke2007hybrid}. 
In particular, Burke's 1999 paper \cite{burke1999integrating} is the first to discuss specifically using these 
two techniques in the same cascaded style that I implement here.

@@

\subsection{Data preparation and feature selection}
The dataset is provided in JSON format, but in such a way that every entry is its own JSON on its own line, 
which does not lend well to fast lookups by the main program. 
Therefore it was necessary to convert every file I used in the system into a typical JSON file, such that it could 
be loaded as a python dictionary with the lookup key being either the uniquely identifying $business\_id$ or 
$user\_id$ depending on the case. 
This allows for extremely fast data lookups in the system. 
It was of course also necessary to filter the list of all Yelp businesses down to just the restaurants. 
I further narrowed the list of restaurants by selecting only those which fit into one or more of the top 12 
categories in terms of number of restaurants (American, Mexican, Italian, Chinese, Seafood, Japanese, Canadian, 
Mediterranean, Indian, Thai, Middle Eastern, Vietnamese, in that order). 
Since the user selects one of these 12 options for categories of restaurants they want to find, they will 
never be recommended a restaurant which doesn't fit at least one of these categories. 
I chose to use the top 12 categories because it's enough to provide a wide variety of options to the user without 
being so many that the user is overwhelmed by the amount. 
I also ensured that every restaurant was still open and had at least 1 check-in in 2019 (the last year that 
data was available). 
This left 20,565 restaurants which could be recommended to the user, which needed to be split into their respective 
metropolitan areas by using the latitude and longitude values of the businesses and of the centres of each city. 
All other datasets were filtered to only data pertaining to these restaurants, and the reviews were further filtered 
to only those which had 5 or more 'useful' votes by other users. 
This was to ensure I was only using high quality reviews which others found helpful, as well as to allow the 
recommender system to compute results in a reasonable amount of time (~10 seconds on my machine). 
This left around 120,000 reviews, which were then converted to a CSV file ($reviews.csv$) and structured as follows: 
($User\_ID$, $Business\_ID$, $Rating$).

20,565 had at least one check-in in 2019, but 333 of these were temporarily closed due to covid but should be 
open now, and 1 is still closed until June 2021. 
19,862 restaurants had a review written about them in 2019, there were 503,979 reviews written in 2019 about these restaurants. 
There are 2,956,316 reviews of these restaurants. 
481,073 tips for these restaurants. 
1,968,703 users. 
None of these restaurants are temporarily closed according to the covid features data. 
20,573 restaurants listed with covid features, 4 restaurants which make up 8 duplicate entries 
(4 have 3 entries), the only difference between their entries being in the 
"Virtual Services Offered" section. 
Allow to pick user ID of 10 users with most reviews made, to avoid cold start problem
4 users have over 10,000 reviews, the most is 14,455 (Victor, 8k3aO-mPeyhbR5HUucA5aA). 
I actually want to select the most reviews of the places I am accepting, restaurants. 
When taking the most reviews of restaurants in teh dataset, a threshold of 600 reviews gives a list of 10 users, 
with the most being 1747. 
I then checked each of these for how many different restaurants they had reviewed and the greatest number of times 
they had reviewed a single restaurant to check that none of them were just users who reviewed a small number of 
restaurants a huge amount of times each. 
All 10 users passed this criteria, with even the least active user having reviewed 496 unique restaurants, which will 
certainly avoid any cold-start problems. 
I split the businesses into 10 groups, by the city which they are closest to, I did this by using latitude and 
longitude rather than State or City since this does not give actual distance and restaurants close to a border 
could then be mis-categorised. 
Since there are 4,223,201 reviews of restaurants I filtered them by removing any which had zero 'useful' votes. 
When filtered to reviews with at least 1 'useful' vote, there are 1,735,732 and when at least 2 votes, 836,929
At least 3 votes, 462,412, at least 5 votes 190,902. 
51,737 with at least 10 votes

Top categories: American, Mexican, Italian, Chinese, Seafood, Japanese, Canadian, Mediterranean, Indian, Thai, Middle Eastern, Vietnamese

The entire Yelp dataset is huge and much of it is not necessary for my domain of bars.
As such I prepared the data by eliminating any data not relevant to my domain.
I selected features such as user ratings and a particular user's average rating,
since if a user typically rates places they go highly, but rates a particular bar low,
this is more significant than a user who always rates places low.

\subsection{Hybrid scheme}
Which two algorithms
A hybrid scheme is a good way to design a recommender system, since you can get the
best of both algorithms if done properly. 
I am using a Cascade Hybrid Recommender System, since this has been shown to accurate predictions 
in my domain of restaurant recommendations \cite{burke2007hybrid}. 
The two systems are Collaborative Filtering and Content Based.

Meaning better recommendations than either algorithm could achieve individually.
Cascade with knowledge-based and collaborative

\subsection{Recommendation techniques/algorithms}
The first recommender system is collaborative filtering. reference https://surpriselib.com/
I am using item-based collaborative filtering, since the number of users is larger than the number of items 
(1,968,703 users vs 209,393 businesses in the entire dataset), this gives greater accuracy of predictions. 
In this case the list of available items does not change, so a user-based method would be unnecessary. 
One of the requirements for this system is justifiability, and item-based methods make it much easier to 
explain why certain predictions were made. 
Item-item collaborative filtering looks for restaurants that are similar, in terms of how people rate them, 
to the restaurants that the user has already rated and recommend the most similar restaurants. 
Filtering reviews by city and then using CF won't work because the active user may not have ever reviewed a restaurant 
in that city, so there is no way to assess similar restaurants to ones they have already rated. 
Instead I carried out CF on all high quality reviews and produce an ordering of recommendations, which I then filter 
by city. 
Mention Netflix Prize and proposition of SVD, use lecture slides explanation and surprise documentation.

I am using a weighted mixed combination of these two systems in order to produce the results
of my hybrid recommender system.

@@

\subsection{Recommendation algorithm}
The idea is to find restaurants in a new city which are similar to those in another city that the user likes, just like Entree. 
I used a model-based approach to Collaborative Filtering based on matrix factorisation. 
The specific model is Single Value Decomposition (SVD), which was popularised by Simon Funk during the Netflix Prize. 
In my program, I fit the algorithm on a training set, and then for every restaurant of the specified category 
and in the specified city, I predict what the user would rate that restaurant. 
Then I sort those restaurants by their score and take the top 8 to present to the user.

\subsection{Output interface}
Recommendations and prediction scores are presented to the user via the command line interface and using an ASCII table. 
I used a python module called texttable for this, since it makes the output very clear and visually appealing. 
I chose to output the top 8 recommendations since I think this is enough variety and choice for the user to find 
a restaurant to try, without there being too many options which might overwhelm the user. 
If I presented more than 8 recommendations at a time, the table wouldn't fit in a single terminal output window and 
would require the user to scroll around to read different parts of the output. 
Since the user will be interacting with the system on a laptop or desktop computer through a command line interface, 
an ASCII table is the most visually attractive way to present information, and the white text on a (typically) black 
background will help accommodate users who are colourblind or who require a high-contrast interface. 
The interface is clear enough to understand that users of all ages are able to use the system, and if a user 
requires the text to be larger then they can increase the text size of their terminal. 
Along with the recommended restaurants, an explanation of how these results are calculated is also provided, as 
well as the prediction values themselves so the user can understand why they are being shown these suggestions. 
This, along with an option in the main menu to view information about the data the system uses and how it is used, 
further adds to the explainability and transparency.

@@

\subsection{Further developments}
What could I do in the future \cite{5284958}. 
Implement the idea of particular categories being closer some than others. 
Specify price range and/or consider price when comparing similarities of restaurants.

>>>>

\subsection{Related work review}
Hybrid recommender systems which combine knowledge-based (KB) and collaborative-filtering (CF) techniques 
are designed to combine the ability of KB techniques to give generic recommendations to new users with 
small profiles and the strength of CF to find peer users with unexpected shared preferences. 
This combination of techniques is particularly well suited to restaurants since a user may often be in the 
mood for a specific type of food, such as Italian or Chinese, so they only want to see suggestions of that 
type or restaurant, which is where the knowledge-based recommender is helpful. 
A collaborative-filtering algorithm will then be able to order all the restaurants which match this 
explicit criteria by how the user should rate them, based on the opinions of other similar users, 
helping them to discover new restaurants to try. 
These techniques are used in the Entree restaurant recommender system \cite{burke1999integrating,burke2000knowledge,burke2002hybrid,burke2007hybrid}. 
In particular, Burke's 1999 paper \cite{burke1999integrating} is the first to discuss specifically using these 
two techniques in the same cascaded style that I implement here.

@@

\subsection{Data preparation and feature selection}
The entire Yelp dataset contains a lot of data which is not relevant to my domain of restaurants, so much data 
preparation is required to save computation time and in order to provide more accurate predictions. 
The dataset is provided in JSON format, but in such a way that every entry is its own JSON on its own line, 
which does not lend well to fast lookups by the main program. 
Therefore it was necessary to convert every file I used in the system into a typical JSON file, such that it could 
be loaded as a python dictionary with the lookup key being either the uniquely identifying $business\_id$ or 
$user\_id$ depending on the case. 
This allows for extremely fast data lookups in the system. 
It was of course also necessary to filter the list of all Yelp businesses down to just the restaurants. 
I further narrowed the list of restaurants by selecting only those which fit into one or more of the top 12 
categories in terms of number of restaurants (American, Mexican, Italian, Chinese, Seafood, Japanese, Canadian, 
Mediterranean, Indian, Thai, Middle Eastern, Vietnamese, in that order). 
Since the user selects one of these 12 options for categories of restaurants they want to find, they will 
never be recommended a restaurant which doesn't fit at least one of these categories. 
I chose to use the top 12 categories because it's enough to provide a wide variety of options to the user without 
being so many that the user is overwhelmed by the amount. 
I also ensured that every restaurant was still open and had at least 1 check-in in 2019 (the last year that 
data was available). 
This left 20,565 restaurants which could be recommended to the user, which needed to be split into their respective 
metropolitan areas by using the latitude and longitude values of the businesses and of the centres of each city. 
All other datasets were filtered to only data pertaining to these restaurants, with duplicates in the 
covid dataset removed. 
The reviews were further filtered to only those which had 5 or more 'useful' votes by other users. 
This was to ensure I was only using high quality reviews which others found helpful, as well as to allow the 
recommender system to compute results in a reasonable amount of time (~10 seconds on my machine). 
This left around 120,000 reviews, which were then converted to a CSV file ($reviews.csv$) with unnecessary data 
removed and structured as follows: ($User\_ID$, $Business\_ID$, $Rating$).

\subsection{Hybrid scheme}
The two personalised recommender systems I implement are knowledge-based and collaborative-filtering, and 
the hybrid scheme is cascade. 
A cascade hybrid scheme works particularly well in my domain of restaurants with an emphasis on providing accurate 
predictions of restaurants which a user may like in a city they have never been to before, based on the restaurants 
they like in their home-city. 
This is because the knowledge-based system can filter the restaurants by location and by the type of food they serve, 
while the collaborative-filtering system is able to find restaurants which are similar to the ones the user has 
previously rated highly. 
So when these two systems are combined in a cascade style with the knowledge-based system as the primary recommender 
and the collaborative-filtering system as the secondary the hybrid system can find hidden similarities between 
restaurants in different cities very accurately. 
This method of hybrid recommender system (HRS) has been implemented as a restaurant recommender in the Entree 
system, giving very accurate results \cite{burke1999integrating,burke2000knowledge,burke2002hybrid,burke2007hybrid}.

\subsection{Recommendation techniques/algorithms}
The first prediction method used in this hybrid system is a knowledge-based (KB) approach, which relies on knowledge 
about each of the restaurants based on its location and the type of food that it serves. 
This approach completely eliminates the cold-start problem, a new user will still receive relevant recommendations 
for the city they are in and the type of food they in the mood for when using a KB system. 
The second prediction method is a collaborative-filtering (CF) approach, specifically Single Value Decomposition (SVD) 
matrix factorisation. 
This approach was chosen because it scales well with large datasets, which is important since my dataset has 
~120,000 entries. 
The CF technique here also introduces an element of variability in the recommendations, such that the user 
may not receive the same ordering of suggestions when they run the system multiple times with the same 
KB specifications, helping the user to try new things. 
This system allows the user to make ratings through the user interface, providing the ability to experiment with 
the system in terms of how new ratings affect the current user's recommendations as well as those of other similar users.

@@

\subsection{Recommendation algorithm}
The collaborative-filtering aspect of the algorithm is carried out by first converting the ratings data to a 
matrix with $User\_ID$ as the rows and $Business\_ID$ as the columns. 
If a user has not rated a restaurant then the rating in that cell is set to 0. 
The prediction $\widehat{r}_{ui}$ (user $u$'s rating of item $i$) is set as:
\begin{equation}
    \widehat{r}_{ui} = \mu + b_u + b_i + q_i^Tp_u
\end{equation}
Where the parameters are learnt by minimising the following regularised squared error: 
\begin{equation}
    \sum_{r_{ui}\in{R_{train}}} (r_{ui} - \widehat{r}_{ui})^2 + \lambda(b_i^2 + b_u^2 + ||q_i||^2 + ||p_u||^2)
\end{equation}
This minimisation is performed by stochastic gradient descent:
\begin{align}
    \begin{split}
    & b_u \gets b_u + \gamma(e_{ui} - \lambda b_u) \\
    & b_i \gets b_i + \gamma(e_{ui} - \lambda b_i) \\ 
    & p_u \gets p_u + \gamma(e_{ui} \cdot q_i - \lambda p_u) \\ 
    & q_i \gets q_i + \gamma(e_{ui} \cdot p_u - \lambda q_i) \\
    \end{split}
\end{align}
where $e_{ui} = r_{ui} - \widehat{r}_{ui}$. 
These steps are performed over all the ratings of the trainset and repeated 20 times \cite{SVD}.

The idea is to find restaurants in a new city which are similar to those in another city that the user likes, just like Entree. 
I used a model-based approach to collaborative-filtering based on matrix factorisation. 
The specific model is Single Value Decomposition (SVD), which was popularised by Simon Funk during the Netflix Prize. 
In my program, I fit the algorithm on a training set, and then for every restaurant of the specified category 
and in the specified city, I predict what the user would rate that restaurant. 
Then I sort those restaurants by their score and take the top 8 to present to the user.

\subsection{Output interface}
Recommendations and prediction scores are presented to the user via the command line interface and using an ASCII table. 
I used a python module called texttable for this, since it makes the output very clear and visually appealing. 
I chose to output the top 8 recommendations since I think this is enough variety and choice for the user to find 
a restaurant to try, without there being too many options which might overwhelm the user. 
If I presented more than 8 recommendations at a time, the table wouldn't fit in a single terminal output window and 
would require the user to scroll around to read different parts of the output. 
Since the user will be interacting with the system on a laptop or desktop computer through a command line interface, 
an ASCII table is the most visually attractive way to present information, and the white text on a (typically) black 
background will help accommodate users who are colourblind or who require a high-contrast interface. 
The interface is clear enough to understand that users of all ages are able to use the system, and if a user 
requires the text to be larger then they can increase the text size of their terminal. 
Along with the recommended restaurants, an explanation of how these results are calculated is also provided, as 
well as the prediction values themselves so the user can understand why they are being shown these suggestions. 
This, along with an option in the main menu to view information about the data the system uses and how it is used, 
further adds to the explainability and transparency. 
\begin{figure}
    \centering
    \includegraphics[width=0.5\linewidth]{output.png}
    \caption{System output}
    \label{fig:output}
\end{figure}

@@

\subsection{Further developments}
What could I do in the future \cite{5284958}. 
Implement the idea of particular categories being closer some than others. 
Specify price range and/or consider price when comparing similarities of restaurants. 
Include demographic information
